---
title: "IE360 Homework 2"
author: "Ugur Kumcu"
date: '2022-05-08'
output: html_document
---


## 1 - INTRODUCTION

  In this Homework, first of all, data will be sketched and then I am going to analyze the data using regression models. After all the analyzes done I will forecast the upcoming 4 quarters which represents the 2007.

### Used Packages
```{r message=FALSE, warning=FALSE}
library(zoo,warn.conflicts = FALSE)
library(readxl)
library(ggplot2)
library(lubridate)
library(corrplot)
require(data.table)
require(GGally)
require(forecast)
```

Reading Data

```{r,include=TRUE, warning=FALSE, error=FALSE}
data_path <- "IE360_Spring22_HW2_data.csv"
data <- read.csv(data_path)
data$Unleaded.Gasoline.Sale..UGS. = gsub(" ", "", data$Unleaded.Gasoline.Sale..UGS.)
data$Unleaded.Gasoline.Sale..UGS. = as.numeric(data$Unleaded.Gasoline.Sale..UGS.)
data$X..LPG.Vehicles..NLPG. = gsub(" ", "", data$X..LPG.Vehicles..NLPG.)
data$X..LPG.Vehicles..NLPG. = as.numeric(data$X..LPG.Vehicles..NLPG)
data$X..Unleaded.Gasoline.Vehicles..NUGV. = gsub(" ", "", data$X..Unleaded.Gasoline.Vehicles..NUGV.)
data$X..Unleaded.Gasoline.Vehicles..NUGV. = as.numeric(data$X..Unleaded.Gasoline.Vehicles..NUGV.)
data$GNP.Agriculture = gsub(" ", "", data$GNP.Agriculture)
data$GNP.Agriculture = as.numeric(data$GNP.Agriculture)
data$GNP.Commerce = gsub(" ", "", data$GNP.Commerce)
data$GNP.Commerce = as.numeric(data$GNP.Commerce)
data$GNP.Total = gsub(" ", "", data$GNP.Total)
data$GNP.Total = as.numeric(data$GNP.Total)
data$Quarter = as.character(data$Quarter)
data = data.table(data)
head(data,8)
##First 2 years data
```

## 2 - Plotting & Commenting

```{r,include=TRUE, warning=FALSE, error=FALSE}
data$UGS = ts(data$Unleaded.Gasoline.Sale..UGS., start = 2000, deltat= 1/4)
ggplot(data, aes(Quarter,data$Unleaded.Gasoline.Sale..UGS. , group=1)) + geom_line(color="red") +geom_point()+ ggtitle("UGS 2000-2006")
##Basic plotting 
```
Basically, we can see decreasing trend and seasonality in the data. Also, variance is getting smaller at the end of the data. Because of these reasons this data is not stationary.
```{r}
acf(data$Unleaded.Gasoline.Sale..UGS.[0:28], 10, main="ACF up to 10 LAGS")
## AutoCorrelation..
```
It can be seen that there is an autocorrelation in lag 4 which can be interpreted as there is an seasonality in the data.
```{r}
data[ , trend := 1 :.N]
## Adding Trend to the Data
seasonality_period <- 4
data[ , seasonality := (1:.N) %% seasonality_period]  
## Adding Seasonality to the Data.  
data[seasonality == 0, seasonality := 4]
## Arranging numbers to see the seasonality
data$seasonality = as.factor(data$seasonality) 
summary(data)
```

## 3 - Regression Models

I am going to improve my model from scratch using other related variables in the data.
```{r}
Model_0 = lm(Unleaded.Gasoline.Sale..UGS.~trend ,data)
## Initial base model
checkresiduals(Model_0$residuals)
summary(Model_0)
```
We need to add seasonality to the model
```{r}
Model_1 = lm(Unleaded.Gasoline.Sale..UGS.~trend + seasonality ,data)
checkresiduals(Model_1$residuals)
summary(Model_1)
```
Model_1's Autocorrelation seems better than the Model_0
Now my main aim is to increase the Adjusted R square value using independent variables.
```{r}
Model_2 = lm(Unleaded.Gasoline.Sale..UGS.~trend + 
            seasonality +
            X..of.Diesel.Gasoline.Vehicles..NDGV. +
            Price.of.Diesel.Gasoline..PG. +
            X..Unleaded.Gasoline.Vehicles..NUGV.,
data)
checkresiduals(Model_2$residuals)
summary(Model_2)
```
Just a small improvement in Adjusted R square value.
```{r, warning=FALSE}
setDT(data)
data[,.(Lag_1 = shift(data$Unleaded.Gasoline.Sale..UGS.,fill=NA, n=1L))]
data$Lag_1 = data[,.(Lag_2 = shift(data$Unleaded.Gasoline.Sale..UGS., fill=NA, n=1L))]
Model_3 = lm(Unleaded.Gasoline.Sale..UGS.~trend + 
            seasonality +
            X..of.Diesel.Gasoline.Vehicles..NDGV. +
            Price.of.Diesel.Gasoline..PG. +
            X..Unleaded.Gasoline.Vehicles..NUGV. +
            Lag_1,
data)
checkresiduals(Model_3$residuals)
summary(Model_3)
```
Lagged variable is also added to data and improved our results.
One can conclude that the Model_3 is better than the previous models. So my predictions will be based on the Model_3.

## 4 - Predict year 2007

Predicting..
```{r}
copy_data = copy(data)
copy_data[,actual:=Unleaded.Gasoline.Sale..UGS.]
copy_data[,prediction:=predict(Model_3,copy_data)]
print(copy_data)
copy_data[ 30, "Lag_1"] = 666210.0
copy_data[,predicted:=predict(Model_3,copy_data)]
print(copy_data)
copy_data[ 31, "Lag_1"] = 840163.9
copy_data[,predicted:=predict(Model_3,copy_data)]
print(copy_data)
copy_data[ 32, "Lag_1"] = 970373.9
copy_data[,predicted:=predict(Model_3,copy_data)]

```

Plotting

```{r, warning=FALSE}
ggplot(copy_data ,aes(x=Quarter)) +
geom_line(aes(y=actual,color='Real_Data', group = 1)) + 
geom_line(aes(y=predicted, color = 'My_Prediction', group = 1) ) 
print(copy_data$predicted)

```

## 5 - Conclusion

I have sketched the main data to have a overall perspective. I have realized that the data has a decreasing trend and a seasonality. Therefore, I have concluded that this data is not stationary due to definition of the stationary. After that, I have created a regression model to predict the upcoming year's data using different techniques which are in "Regression Models" section. Lastly, I have sketched all of my predictions about 2007.  
